\chapter*{Acknowledgements}
\addcontentsline{toc}{chapter}{Acknowledgements}
{\bfseries Lenny}, I could not have wished for a better mentor. Coming to you with whatever problem I had invariably made the problem less bad. Your relentless optimism is painfully infective, and I hope to carry it for the rest of my life. \\

{\bfseries Brendan}, I made a directed graph of the people in these acknowledgements, the edges signifying who I met through whom. Your efforts to socialize the department put you squarely at the top of the largest connected component of the graph. Thank you for the excellent boating, climbing, brewery-touring, and thank you for being the root of my digraph.

{\bfseries H\'el\`ene}, you've taken the Allez pas les bleus people from France to Spain to exotic Friesland, and hopefully some day to Font. You've worked at giving us all hot yoga bodies, while at the same time feeding us wine. Thanks for pointing out the bright side when I feel like there isn't one, and thanks for making Allez pas what it is.

{\bfseries Lorenzo}, from organizing cinema seminars to filling buckets with rice not to feed hungry kids in Africa, but to just, you know, kind of squeeze the rice every now and then, you are the key reason why working at the KdVI was as fun as it was. Thanks for the climbing, the limoncello, the dinner parties, for choking me during my first (and absolutely final) bjj class I took, and for the fun lunches.

{\bfseries Madi}, I originally wanted to make your entire acknowledgement out of puns, but it was much harder than expected. I just couldn't quite fit them in. Thank you for all the movies, the chats, and for making me understand that life is pain au chocolat.

{\bfseries Sam}, thank you for allowing me to take shelter in the comfort of your various toilets.

{\bfseries Sarah}, thank you for all the chats, little cakes, tea parties, running, yoga, swimming, and all the classes we took on that health and fitness app which I probably can't mention by name because it could possibly be construed as advertising. If you ever find yourself back in Amsterdam I hope we can finally do an aqua swing class. I'll try and get Henry to join, I'm sure he misses you.

{\bfseries Thibault}, thanks for being one of the main reasons conferences were fun to go to, and for providing me with a home away from home in Leiden. Here's to Fran\c{c}ois. \\

Many thanks go out to the people at the KdVI, for making my time there so enjoyable. I'm particularly thankful to {\bfseries Beau Darius, Drugo, Gerald, Mulva, Omniscientis, Rhinoa, and Vorondil}, and especially {\bfseries Fabian} for keeping them in check.

{\bfseries Arie, Dani, Jeroen, Kirsten, Reinier, Renjie, Sasha, Zijian, Zo\"e}, thank you for picking up the night and weekend shifts with me, the dinners, the interesting lunches, the laughs, the shared birthday party, the typographical and linguistic discussions, the gruesome viking mud run, and thank you for putting up with the fire hazard that was my desk. In my heart, we will always be office mates.

{{\bfseries Mees}, at the very start of my existence as a PhD student, you baked some brownies. They were lovely. Thank you.}

{\bfseries Bart, Jimme, Karin}, thanks to you it is so natural to me that families are supportive, accepting, and forgiving to a fault, that I sometimes forget it's not a given. Thank you for providing me with all of those things.

{\bfseries Sujin}, 흰둥이!

{\bfseries Max}, you called me up just to ask how I was doing, you gave me melatonin when you found out I was sleeping poorly, you gave me tips on vaginal hygiene. You were the sweetest big little brother. I would give anything for another chance at having you beat me up with a giant inflatable chanticleer. Actually, I mainly wish you were still alive so we all wouldn't miss out on so much of you, and that I would've been around you more when you were alive. The chanticleer thing is kind of secondary to that. Even though, as we both know, there is no such thing as an afterlife, picturing you looking down on us smugly from the danger zone while having us listen to Gerard Joling at your funeral will always make me laugh.
